---
title: "p8105_hw2_ly2496"
author: "Ling Yi"
date: "9/23/2020"
output: github_document
---

Problem 0

##loading the packages needed

```{r}
library(tidyverse)
library(readxl)
```


?read_excel

##problem 1

```{r}
trashwheel_df =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A:N")) %>%
            janitor:: clean_names() %>%
            drop_na(dumpster)%>%
          mutate(
              sports_balls=round(sports_balls),
              sports_balls = as.integer(sports_balls)
          )
```

##read in precipitation data for 2018

```{r}
precip_2018 =
  read_excel ("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
              sheet = "2018 Precipitation",
              skip = 1) %>% 
              janitor:: clean_names() %>%
              drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

##read in precipitation data for 2017

```{r}
precip_2017 =
  read_excel ("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
              sheet = "2017 Precipitation",
              skip = 1) %>% 
              janitor:: clean_names() %>%
              drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```


##combining the precipitation dataset and adding in month name

```{r}
month_df=
  tibble(
    month =  1:12,
    month_name=month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) %>%  view


left_join(precip_df, month_df, by = "month")

```


This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects the trash, and stores it in a dumpster. The dataset contains information per year, month, and trash collected include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` in the final data set. Additional data sheets include month precipitation data. 

The total precipitation for 2018 is `r precip_df %>%  filter(year=="2018") %>%  summarize(sum(total))`(in) and the medium number of sports balls in the dumpter in 2017 is `r trashwheel_df %>% filter (year=="2017") %>% summarize(median(sports_balls))`. 


##problem 2

```{r}
nyc_transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(2:18, "vending", "ada") %>% 
  mutate(entry = 
    recode(entry, "YES"="TRUE", "NO"= "FALSE"))
  
```

##how many distinct stations

```{r}
nyc_transit_df %>% 
  distinct(station_name) %>% 
  count()


##There are 356 disctinct stations
```

##how mnay distinct lines

```{r}
nyc_transit_df %>% 
  distinct(line) %>% 
  count()

##there are 36 disctint subway lines
```


##how many statiosn are ada compliant

```{r}
filter(nyc_transit_df, ada =="TRUE") %>% nrow()

##there are 468 stations that are ada compliant
```

##proportion of entrances/exist without vending

```{r}
nyc_transit_df %>% 
    count(vending, entry) %>% 
    group_by(entry) %>% 
    transmute(vending, Percentage=n/sum(n)*100)
## the proportion of stations with entrances but without vending is 3.93%
## As TA about this
```

##Problem 3

```{r}
month_pol_df = 
  read_csv("./data/problem_3/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    president = ifelse(prez_gop == 1, "gop", 
                       ifelse(prez_dem == 1, "dem", "NA"))
    ) %>% 
  select(-c("prez_dem", "prez_gop", "day"))


  
```







