---
title: "p8105_hw2_ly2496"
author: "Ling Yi"
date: "9/23/2020"
output: github_document
---

Problem 0

##loading the packages needed

```{r}
library(tidyverse)
library(readxl)
```


?read_excel

##problem 1

```{r}
trashwheel_df =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A:N")) %>%
            janitor:: clean_names() %>%
            drop_na(dumpster)%>%
          mutate(
              sports_balls=round(sports_balls),
              sports_balls = as.integer(sports_balls)
          )
```

##read in precipitation data for 2018

```{r}
precip_2018 =
  read_excel ("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
              sheet = "2018 Precipitation",
              skip = 1) %>% 
              janitor:: clean_names() %>%
              drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

##read in precipitation data for 2017

```{r}
precip_2017 =
  read_excel ("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
              sheet = "2017 Precipitation",
              skip = 1) %>% 
              janitor:: clean_names() %>%
              drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```


##combining the precipitation dataset and adding in month name

```{r}
month_df=
  tibble(
    month =  1:12,
    month_name=month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)


left_join(precip_df, month_df, by = "month")

```


This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects the trash, and stores it in a dumpster. The data set contains information per year, month, and trash collected include some specific kinds of trash. The variables included in this data set is `r names(trashwheel_df)`. There are a total of `r nrow(trashwheel_df)` observations in the final data set. The data set has information from `r min(pull(trashwheel_df, year))` to `r max(pull(trashwheel_df, year))`. Additional data sheets include month precipitation data. 

The total precipitation for 2018 is `r precip_df %>%  filter(year=="2018") %>%  summarize(sum(total))`(in) and the median number of sports balls in the dumpster in 2017 is `r trashwheel_df %>% filter (year=="2017") %>% summarize(median(sports_balls))`. 


##problem 2

```{r}
nyc_transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(
    Route8 = col_character(),
    Route9 = col_character(),
    Route10 = col_character(),
    Route11 = col_character()
  )) %>% 
  janitor::clean_names() %>% 
  select(2:18, "vending", "ada") %>% 
  mutate(entry = 
    recode(entry, "YES"="TRUE", "NO"= "FALSE"))

```

##how many distinct stations

There are `r nyc_transit_df %>% distinct(line, station_name) %>% count()` stations in nyc.


##how many distinct lines

There are `r nyc_transit_df %>% distinct(line) %>% count()` subway lines


##how many stations are ada compliant

There are `r filter(nyc_transit_df, ada =="TRUE") %>% distinct(line, station_name) %>% count()` distinct stations and lines that are ada compliant

```{r}

filter(nyc_transit_df, ada =="TRUE") %>% 
  distinct(line, station_name) %>% 
  count()

```


##proportion of entrances/exist without vending

The proportions of stations without vending but has entrances is `r filter(nyc_transit_df, vending == "NO" & entry =="TRUE") %>% count()/ filter(nyc_transit_df, vending == "NO") %>% count()`


##reformatting the route variable

```{r}
  nyc_transit_df =
   pivot_longer(
    nyc_transit_df,
    route1:route11, 
    names_to ="route",
    values_to = "route_name"
  ) %>% 
  separate(route, into = c("route", "route_number"), sep = 5) %>% 
  select (-c("route"))

filter(nyc_transit_df, route_name =="A") %>% distinct(line, station_name) %>% count()
```

There are `r filter(nyc_transit_df, route_name =="A") %>% distinct(line, station_name) %>% count()` distinct stations that serve the A train. 

Of the stations that serve train A, there are `r filter(nyc_transit_df, route_name =="A") %>% filter(ada =="TRUE") %>% distinct(line, station_name) %>%  count()` that are ada compliant. 


##Problem 3

```{r}
month_pol_df = 
  read_csv("./data/problem_3/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    president = ifelse(prez_gop == 1, "gop", 
                       ifelse(prez_dem == 1, "dem", "NA"))
    ) %>% 
  select(-c("prez_dem", "prez_gop", "day")) %>% 
  mutate(
    month = as.integer(month),
    year = as.character(year)
  )


##add in month names
month_pol_df = left_join(month_pol_df, month_df, by = "month") %>% 
  select (-c("month")) %>% 
  relocate ("year", "month_name")

```

##improting snp data set


```{r}
snp_df = 
  read_csv("./data/problem_3/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  select(-c("day")) %>% 
  mutate(
    month=as.integer(month)
)

##add in month names
snp_df = left_join(snp_df, month_df, by = "month") %>% 
  select (-c(month)) %>% 
  relocate ("year", "month_name")


```


```{r}
unemploy_df = 
  read_csv("./data/problem_3/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_pct"
  )%>% 
  mutate (
    Year = as.character(Year)
  )


month_abv_df=
  tibble(
    month_abv =  1:12,
    month=month.abb,
    month_name=month.name
  ) 

unemploy_df = left_join(unemploy_df, month_abv_df, by = "month") %>% 
  select(-c("month_abv", "month")) %>% 
  janitor::clean_names()

```

## joining the datasets

```{r}

combined_df = left_join(month_pol_df, snp_df, by = c("year", "month_name"))
list(unemploy_df)

final_df = left_join(combined_df, unemploy_df, by = c("year", "month_name"))
```


##describing the data set

The data sets are sourced from the FiveThrityEight website and we used three data sets in the code above. The variables in the unemployment dataset are `r names(unemploy_df)`. There is a total of `r nrow(unemploy_df)` observations in this data set. The data set contains information from the year `r min(pull(unemploy_df, year))` to the year `r max(pull(unemploy_df, year))`. Between this time range, the mean percent of unemployment is `r mean(pull(unemploy_df, unemployment_pct), na.rm=TRUE)`. There is a total of `r sum(is.na(unemploy_df))` missing values in this data set. 

The other data set (snp_df) is about stock values. The variables in this data set are `r names(snp_df)` and there is a total of `r nrow(snp_df)` observations in this data set. The variable "close" means the closing values of the S&P stock index on the associated date. The data in the data set ranges from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. The mean closing value for this time period is `r mean(pull(snp_df, close), na.rm=TRUE)` with a minimum of `r min(pull(snp_df, close), na.rm = TRUE)` and a max of `r max(pull(snp_df, close), na.rm = TRUE)`. There is a total of `r sum(is.na(snp_df))` missing values in this data set. 

The third data set (month_pol_df) is about the number of national politicians who are democratic or republican at any given time. The variables in this data set are `r names(month_pol_df)` and there is a total of `r nrow(month_pol_df)` observations in this data set. The data in the data set ranges from `r min(pull(month_pol_df, year))` to `r max(pull(month_pol_df, year))`. There is a total of `r sum(is.na(month_pol_df))` missing values in this data set. 


When combining all three data sets together, the resulting data set consists of `r nrow(final_df)` observations. The variables in this data set are `r names(final_df)` after combining by year and month. The resulting dataset have a total of `r sum(is.na(final_df))` missing values.



